Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-06-06 paper/main.tex

\subsection{The experimental workflow}

We show our DBTL workflow in Figure \ref{fig: Flowchart}.
BUILD and TEST are driven chiefly by choices made by human researchers and use of automated methods.
Machine learning algorithms are applied in LEARN and DESIGN.
In LEARN phase, we use the Gaussian Process regression algorithm to predict the TIR of RBS sequences comprising the experimental space.
The goal is to provide predictions about TIR labels and uncertainty to assist the recommendation process.
In the DESIGN phase, we have used a multi-armed Bandit recommendation algorithm that was selecting RBS designs to be tested in subsequent experimental batches.\\

@@

\subsection{Design of the investigated genetic device}

@@

\subsection{Performance of the recommendation algorithm}

The design recommendations were made using the Multi-armed Bandit algorithm.
In short, this algorithm is a stochastic method of probing of the experimental space. 
This algorithm aims at maximising the reward (output) from testing a limited number of instances from a big pool which cannot be wholly tested due to limited resources (time, computational power, capital). 
In our case we use the Upper Confidence Bound version of the algorithm, which focuses its recommendations on sequences that should give highest TIR based on the probabilities computed by the prediction algorithm (see below). 
Another feature of the bandits algorithm is that it balances two parts: the exploration of the unknown (untested) parts of the design space where high TIR RBS can be hidden, and exploitation which goal is querying areas which are known to give relatively high TIRs.
One thing of note is that the bandit algorithm is stochastic, that is it exploits the probabilities of a given event occurring (in this case RBS having a specific TIR). 
As such, it pairs naturally with our prediction algorithm, the Gaussian Process Regression, which provides probability based function regression.

@@

\subsection{Prediction of RBS performance}

@@

Figure \ref{fig: Scatterplot} shows how our algorithm performed in terms of predictions in each round. 
As expected, the predictions in Round 0 were poor due to use of approximated data. 
The predictions improved for the subsequent rounds, from R\textsuperscript{2} of 0.065 for round 0 to R\textsuperscript{2} of 0.27 for round 3.
Similarly, the Spearman correlation coefficient rose from 0.27 for Round 0 to 0.48 for Round 3.
One important point to note is that the predictions are also influenced by our recommendation choices. 
In each round, we select a number of data points for exploration, which means that these data points, when tested, have a high chance of having real mean different to what was predicted.
However, this is still very useful information for future predictions as it allows us to understand the underlying space better.

@@

\subsection{Characteristics of the tested sequences}

@@

There are still open questions that need to be addressed for applying machine learning in synthetic biology.
For example, how can we understand the language of RBS sequences better?
In other words, can we extract more biologically important information from the decisions made by our algorithms?
Another example would be - given a small amount of RBS sequences tested, how can machine learning algorithm provide more accurate predictions and uncertainty measurement? 
Next question would be - how generalizable is this method? Can it be succesfully used for design of bigger genetic parts, like promoters?
Finally, another question with high impact on the performance of the workflow is - how should we choose the parameters which controls the exploration-exploitation rate? \\

>>>>

We present our RBS optimizing DBTL workflow that uses machine learning in Section~\ref{sec:dbtl-workflow}.
Machine learning (ML) is used in two different ways: we show the efficacy of the ML recommendations
in the DESIGN stage (Section~\ref{sec:ucb-results}),
and we demonstrate that the ML predictions are accurate in the LEARN stage (Section~\ref{sec:gp-results}).
We present our new RBS sequence library in Section~\ref{sec:characteristics-of-library} and decribe some interesting
characteristics of the discovered sequences, 
as well as show the effectiveness of the automated laboratory workflow stages.

\subsection{The experimental workflow}
\label{sec:dbtl-workflow}

\cheng{This section (I merged 2.1 and 2.2), feels like a method section. Rewrite to be a results section. 
This can be much shorter}

We show our DBTL workflow in Figure \ref{fig: Flowchart}, that uses machine learning to optimise protein expression.
BUILD and TEST are driven chiefly by choices made by human researchers and use of automated methods.
Machine learning algorithms are applied in LEARN and DESIGN.
In LEARN phase, we use the Gaussian Process regression algorithm to predict the TIR of RBS sequences comprising the experimental space.
The goal is to provide predictions about TIR labels and uncertainty to assist the recommendation process.
In the DESIGN phase, we have used a multi-armed Bandit recommendation algorithm that was selecting RBS designs to be tested in subsequent experimental batches.\\

@@

%\subsection{Design of the investigated genetic device}

@@

\subsection{DESIGN: Performance of the recommendation algorithm}
\label{sec:ucb-results}

\cheng{Write this section to describe Figure 2, in the order of the subfigures (can rearrange if needed). Again, feels like a method section.}
The design recommendations were made using the Multi-armed Bandit algorithm.
In short, this algorithm is a stochastic method of probing of the experimental space. 
This algorithm aims at maximising the reward (output) from testing a limited number of instances from a big pool which cannot be wholly tested due to limited resources (time, computational power, capital). 
In our case we use the Upper Confidence Bound version of the algorithm, which focuses its recommendations on sequences that should give highest TIR based on the probabilities computed by the prediction algorithm (see below). 
Another feature of the bandits algorithm is that it balances two parts: the exploration of the unknown (untested) parts of the design space where high TIR RBS can be hidden, and exploitation which goal is querying areas which are known to give relatively high TIRs.
One thing of note is that the bandit algorithm is stochastic, that is it exploits the probabilities of a given event occurring (in this case RBS having a specific TIR). 
As such, it pairs naturally with our prediction algorithm, the Gaussian Process Regression, which provides probability based function regression.

@@

\subsection{LEARN: Prediction of RBS performance}
\label{sec:gp-results}

@@

\cheng{Start this section with these results.}
Figure \ref{fig: Scatterplot} shows how our algorithm performed in terms of predictions in each round. 
As expected, the predictions in Round 0 were poor due to use of approximated data. 
The predictions improved for the subsequent rounds, from R\textsuperscript{2} of 0.065 for round 0 to R\textsuperscript{2} of 0.27 for round 3.
Similarly, the Spearman correlation coefficient rose from 0.27 for Round 0 to 0.48 for Round 3.
One important point to note is that the predictions are also influenced by our recommendation choices. 
In each round, we select a number of data points for exploration, which means that these data points, when tested, have a high chance of having real mean different to what was predicted.
However, this is still very useful information for future predictions as it allows us to understand the underlying space better.

@@

\subsection{BUILD+TEST: Characteristics of the tested sequences}
\label{sec:characteristics-of-library}

@@

\cheng{Have a look at the paper I wrote with Mengyan, and summarise the first few opportunities and challenges there.}
There are still open questions that need to be addressed for applying machine learning in synthetic biology.
For example, how can we understand the language of RBS sequences better?
In other words, can we extract more biologically important information from the decisions made by our algorithms?
Another example would be - given a small amount of RBS sequences tested, how can machine learning algorithm provide more accurate predictions and uncertainty measurement? 
Next question would be - how generalizable is this method? Can it be succesfully used for design of bigger genetic parts, like promoters?
Finally, another question with high impact on the performance of the workflow is - how should we choose the parameters which controls the exploration-exploitation rate? \\
